---
title: "Movie and TV animal deaths"
---

## Overview of the analysis pipeline

This notebook carries out the following steps, first for movies and then for TV series:

1.  **Query the Does the Dog Die (DTDD) database.** For each title in the IMDB top 5000 list, query the DTDD API to find a matching entry. Movies are matched by IMDB ID, which is a reliable identifier; TV series are matched by title search, because I found that IDs weren't as reliable for TV series. Title-based matching introduces some risk of false matches, which is addressed later with a deduplication step.

2.  **Filter to valid matches.** Titles with no DTDD match, mismatched release years (to exclude remakes or unrelated titles with the same name), or a small number of ratings (an arbitrary cutoff intended to exclude obscure titles or unreliable matches) are removed. For TV series, an additional deduplication step removes false title-based matches (see inline comments below for details).

3.  **Fetch death vote details.** For each remaining title, query DTDD for the yes/no vote counts on three topics: "a dog dies," "a cat dies," and "an animal dies" (the latter excludes dogs, cats, and horses according to DTDD's trigger definitions).

4.  **Classify deaths by majority vote.** For each title and topic, a death is classified as TRUE if yes votes outnumber no votes, and FALSE otherwise (including in the case of a tie).

5.  **Analyze the classified data.** This includes: (a) computing overall death rate percentages with confidence intervals, (b) testing whether these rates differ significantly between companion animals (dogs, cats) and other animals, (c) examining trends in death rates over time, and (d) examining death rates by genre.

Raw and processed data are saved at each major step (as both `.csv` for human readability and `.rds` for faster reloading) so the pipeline can be resumed without re-querying the API.

## Query the Does the Dog Die database

This script queries the Does the Dog Die? database for movies in the IMDB top 5000 list that are (probably) American and released on or after 2000.

It then repeats this analysis for the top 5000 TV series.

```{r}

# Load libraries

library(httr)
library(jsonlite)
library(tidyverse)
library(janitor)

```

## Helper functions

```{r}

# Function to query the Does The Dog Die database
query_dtdd_database <- function(movie_title) {
  resp <- GET(
    url = "https://www.doesthedogdie.com/dddsearch",
    query = list(q = movie_title),
    add_headers(`Accept` = "application/json", `X-API-KEY` = Sys.getenv("DDD_API_KEY"))
  )
  
  if (status_code(resp) == 200) {
    data <- fromJSON(content(resp, "text", encoding = "UTF-8"), flatten = TRUE)
    if (length(data$items) > 0) {
      return(as_tibble(data$items))
    } else {
      return(tibble(title = movie_title, message = "No results found"))
    }
  } else {
    return(tibble(title = movie_title, message = "Error querying API"))
  }
}

# Function to query Does the Dog Die database by IMDB ID
query_dtdd_by_imdb_id <- function(imdb_id) {
  resp <- GET(
    url = "https://www.doesthedogdie.com/dddsearch",
    query = list(imdb = imdb_id),
    add_headers(`Accept` = "application/json", `X-API-KEY` = Sys.getenv("DDD_API_KEY"))
  )
  
  if (status_code(resp) == 200) {
    data <- fromJSON(content(resp, "text", encoding = "UTF-8"), flatten = TRUE)
    if (length(data$items) > 0) {
      return(as_tibble(data$items))
    } else {
      return(tibble(imdb_id = imdb_id, message = "No results found"))
    }
  } else {
    return(tibble(imdb_id = imdb_id, message = "Error querying API"))
  }
}

# Function to fetch details for a specific title
fetch_dtdd_title_details <- function(item_id) {
  resp <- GET(
    url = paste0("https://www.doesthedogdie.com/media/", item_id),
    add_headers(`Accept` = "application/json", `X-API-KEY` = Sys.getenv("DDD_API_KEY"))
  )
  
  if (status_code(resp) == 200) {
    data <- fromJSON(content(resp, "text", encoding = "UTF-8"), flatten = TRUE)
    # extract the relevant data
    # 1. a dog dies yes votes and no votes
    # To do this, I first identify the topic ID associated with "a dog dies"
    dogVotes <- data$topicItemStats |> filter(tolower(topic.name) == "a dog dies") |> 
      select(topic.name, yesSum, noSum)
    # rename the yes and no columns to denote that they apply only to dogs
    dogVotes <- dogVotes |> 
      rename(dog_yes_votes = yesSum, dog_no_votes = noSum)
    
    # 2. a cat dies yes votes and no votes
    catVotes <- data$topicItemStats |> filter(tolower(topic.name) == "a cat dies") |> 
      select(topic.name, yesSum, noSum)
    # rename the yes and no columns to denote that they apply only to cats
    catVotes <- catVotes |> 
      rename(cat_yes_votes = yesSum, cat_no_votes = noSum)
    
    # 3. an animal dies yes votes and no votes
    animalVotes <- data$topicItemStats |> filter(tolower(topic.name) == "an animal dies") |> 
      select(topic.name, yesSum, noSum)
    # rename the yes and no columns to denote that they apply to all animals
    animalVotes <- animalVotes |> 
      rename(animal_yes_votes = yesSum, animal_no_votes = noSum)
    
    # Join the votes data together with a column for the name of the movie
    data <- tibble(
      dog_yes_votes = dogVotes$dog_yes_votes,
      dog_no_votes = dogVotes$dog_no_votes,
      cat_yes_votes = catVotes$cat_yes_votes,
      cat_no_votes = catVotes$cat_no_votes,
      animal_yes_votes = animalVotes$animal_yes_votes,
      animal_no_votes = animalVotes$animal_no_votes
    )
    
    return(data)
  } else {
    return(tibble(message = "Error fetching details"))
  }
}
```

## Top 5000 movies

```{r}
# Read in the IMDB 5000 movies data set
# Source: https://www.kaggle.com/datasets/tiagoadrianunes/imdb-top-5000-movies
imdb_movies_5000 <- read_csv("data/imdb_movies_5000.csv")

```

### Query the database and pull the data

```{r}
# For each movie in the set of recent movies, query the Does The Dog Die database
# and pull out info from the movie. 
# 
# If the movie doesn't exist in the database (or there are no results), skip it.
# Put the data from all movies into one tibble.


results_all_movies <- imdb_movies_5000 |>
  #slice_head(n = 50) |> # NOTE: This line is for debugging purposes only
  mutate(dtdd_data = map(tconst, query_dtdd_by_imdb_id)) |>
  unnest(dtdd_data) |> 
  # Filter out movies that don't have an ID in the database
  filter(!is.na(id)) |>
  # Filter out movies that don't have at least 100 ratings (arbitrary cutoff)
  # The idea here is that shows with few ratings either have unreliable stats
  # or are obscure or not valid matches
  filter(numRatings > 100) |>
  # Filter out movies that don't match years to eliminate movies that have been
  # remade and/or duplicates
  filter(startYear == releaseYear)



```

```{r}
# Save the results
write_csv(results_all_movies, "data/imdb_5000_dtdd_results.csv") # human readable
saveRDS(results_all_movies, "data/imdb_5000_dtdd_results.rds") # for data processing
```

```{r}
# Fetch the movie details for each movie in the results using the id
# and save all the results into a new tibble

results_movie_details <- results_all_movies |>
  #slice_head(n = 10) |> # NOTE: This line is for debugging purposes only
  mutate(dtdd_details = map(id, fetch_dtdd_title_details)) |>
  unnest(dtdd_details) |>
  select(id, name, cleanName, 
         dog_yes_votes, dog_no_votes,
         cat_yes_votes, cat_no_votes,
         animal_yes_votes, animal_no_votes)
```

```{r}
# Save the movie details
write_csv(results_movie_details, "data/imdb_5000_dtdd_movie_details.csv")
```

### Analyze the data

```{r}
# Create new columns in results_movie_details that calculate 
# the majority vote for each animal type and set to 1 if
# the majority vote is "yes" and 0 if the majority vote is "no".
# Set the value to 0 if there is no majority vote (i.e. if the yes and no votes are equal).
results_movie_details <- results_movie_details |>
  mutate(dog_death = case_when(
    dog_yes_votes > dog_no_votes ~ TRUE,
    TRUE ~ FALSE),
    cat_death = case_when(
      cat_yes_votes > cat_no_votes ~ TRUE,
      TRUE ~ FALSE),
    animal_death = case_when(
      animal_yes_votes > animal_no_votes ~ TRUE,
      TRUE ~ FALSE)
  )

# Save the results for later reference
saveRDS(results_movie_details, "data/imdb_5000_dtdd_movie_details.rds")
```

#### Compute death percentages

```{r}
# Compute percentages with CIs
compute_proportion_ci <- function(data, var) {
  n <- nrow(data)
  x <- sum(data[[var]], na.rm = TRUE)
  test <- prop.test(x, n)
  tibble(
    variable = var,
    pct = test$estimate * 100,
    ci_low = test$conf.int[1] * 100,
    ci_high = test$conf.int[2] * 100
  )
}

# Compute percentages and CIs for movie deaths
movies_proportions <- map_df(
  c("dog_death", "cat_death", "animal_death"),
  ~ compute_proportion_ci(results_movie_details, .x)
)
print(movies_proportions)

# Test whether these proportion differences are statistically significant

# Helper function to run prop.test comparing two proportions
compare_proportions <- function(data, var1, var2) {
  n <- nrow(data)
  x1 <- sum(data[[var1]], na.rm = TRUE)
  x2 <- sum(data[[var2]], na.rm = TRUE)
  prop.test(c(x1, x2), c(n, n))
}

# Compare animal vs dog and animal vs cat
test_animal_vs_dog <- compare_proportions(results_movie_details, "animal_death", "dog_death")
test_animal_vs_cat <- compare_proportions(results_movie_details, "animal_death", "cat_death")

print(test_animal_vs_dog)
print(test_animal_vs_cat)
```

#### Compute correlations between animal categories

```{r}
# Compute the correlation between a movie having a dog death and
# having an animal death
cor.test(as.numeric(results_movie_details$dog_death),
         as.numeric(results_movie_details$animal_death),
         method = "pearson")

# Compute the correlation between a movie having a cat death and
# having an animal death
cor.test(as.numeric(results_movie_details$cat_death),
         as.numeric(results_movie_details$animal_death),
         method = "pearson")

# Compute the correlation between a movie having a dog death and
# having a cat death
cor.test(as.numeric(results_movie_details$dog_death),
         as.numeric(results_movie_details$cat_death),
         method = "pearson")
```

#### Analyze time trends

```{r}
# Generate a plot of animal deaths over time

library(ggrepel)
library(patchwork)

p1 <- results_movie_details |>
  left_join(results_all_movies |> select(id, startYear), by = "id") |>
  filter(startYear >= 1999) |>
  pivot_longer(cols = c(dog_death, cat_death, animal_death),
               names_to = "death_type", values_to = "died") |>
  group_by(startYear, death_type) |>
  summarise(proportion = mean(died, na.rm = TRUE), .groups = "drop") |>
  mutate(death_type = recode(death_type,
    dog_death = "Dog",
    cat_death = "Cat",
    animal_death = "Other"
  )) |>
  group_by(death_type) |>
  mutate(is_last = startYear == max(startYear)) |>
  ungroup() |>
  ggplot(aes(x = startYear, y = proportion, 
             color = death_type, shape = death_type)) +
  geom_point() +
  geom_smooth(alpha = 0.3) +
  annotate("text", 
           x = 2025.5, 
           y = c(0.015, -0.009, 0.26), 
           label = c("Cat", "Dog", "Other"),
           color = "black",
           hjust = 0) +
  scale_color_brewer(palette = "Dark2") +
  scale_shape_manual(values = c(
    "Dog" = 16,
    "Cat" = 17,
    "Other" = 15
  )) +
  scale_y_continuous(labels = scales::percent) +
  scale_x_continuous(limits = c(1998.5, 2027)) +
  labs(x = "", y = "Proportion of\nmovie deaths") +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.title.y = element_text(angle = 0, vjust = 0.5),
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank()
  )

p2 <- results_movie_details |>
  left_join(results_all_movies |> select(id, startYear), by = "id") |>
  filter(startYear >= 1999) |>
  distinct(id, startYear) |>
  count(startYear) |>
  ggplot(aes(x = startYear, y = n)) +
  geom_col() +
  scale_x_continuous(limits = c(1998.5, 2027)) +
  labs(x = "Release year", y = "Number of\nmovies") +
  theme_minimal() +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    axis.title.y = element_text(angle = 0, vjust = 0.5),
  )

ggsave("figures/movies_trend_analysis.pdf", 
       plot = p1 / p2 + plot_layout(heights = c(4, 1), axes = "collect"),
       width = 7, 
       height = 4, 
       units = "in")
```

#### Analyze by genre

```{r}
# Analyze movies by genre
# NOTE: I'm using the genres assigned by IMDB here and there are multiple
# genres for each movie, so this will produce more data points than movies

movies_with_genre <- left_join(results_movie_details, results_all_movies |> select(id, genres), by = "id") |>
  separate_rows(genres, sep = ", ") 

# Pull only the top 15 most common genres
valid_genres <- movies_with_genre |> 
  count(genres) |> 
  slice_max(order_by = n, n = 15) |> 
  # filter(n >= 100) |> 
  pull(genres)

movies_with_genre <- movies_with_genre |> 
  filter(genres %in% valid_genres)

# Calculate death rates by genre and CIs
genre_death_rates <- movies_with_genre |>
  group_by(genres) |>
  summarise(
    n = n(),
    dog_death_rate = mean(dog_death, na.rm = TRUE),
    cat_death_rate = mean(cat_death, na.rm = TRUE),
    animal_death_rate = mean(animal_death, na.rm = TRUE),
    dog_se = sqrt(dog_death_rate * (1 - dog_death_rate) / n),
    cat_se = sqrt(cat_death_rate * (1 - cat_death_rate) / n),
    animal_se = sqrt(animal_death_rate * (1 - animal_death_rate) / n)
  )

# Get genre order based on animal death rate
genre_order <- genre_death_rates |>
  arrange(animal_death_rate) |>
  pull(genres)

label_order <- genre_death_rates |>
  arrange(animal_death_rate) |>
  mutate(genre_label = paste0(genres, " (", n, ")")) |>
  pull(genre_label)

# Pivot to long format
genre_death_rates_long_movies <- genre_death_rates |>
  pivot_longer(
    cols = c(dog_death_rate, cat_death_rate, animal_death_rate),
    names_to = "death_type",
    values_to = "proportion"
  ) |>
  mutate(
    genre_label = factor(paste0(genres, " (", n, ")"), levels = label_order),
    se = case_when(
      death_type == "dog_death_rate" ~ dog_se,
      death_type == "cat_death_rate" ~ cat_se,
      death_type == "animal_death_rate" ~ animal_se
    ),
    death_type = factor(death_type, 
                        levels = c("animal_death_rate", "dog_death_rate", "cat_death_rate"),
                        labels = c("Other", "Dog", "Cat"))
  )
```

## Top 5000 TV series

```{r}
# Read in the IMDB 5000 TV series data set
# Source: https://www.kaggle.com/datasets/tiagoadrianunes/imdb-top-5000-tv-shows
imdb_tvseries_5000 <- read_csv("data/imdb_tvseries_5000.csv")
```

### Query the database

```{r}
# For each series in the set of 5000 top TV series, query the Does The Dog Die database
# and pull out info from the series. 
# 
# If the series doesn't exist in the database (or there are no results), skip it.
# Put the data from all series into one tibble.
#
# Note: I used a title search for TV series rather than an ID search because the ID
# search resulted in too many false and failed matches

results_all_tvseries <- imdb_tvseries_5000 |>
  #slice_head(n = 50) |> # NOTE: This line is for debugging purposes only
  # We discard the genre information from DTDD because the column conflicts with
  # the column from IMDB and we will use that genre info anyway
  mutate(dtdd_data = map(primaryTitle, ~ query_dtdd_database(.x) |> select(-genres))) |>
  #mutate(dtdd_data = map(primaryTitle, query_dtdd_database)) |>
  unnest(dtdd_data) |>
  # Filter out shows that don't have at least 100 ratings (arbitrary cutoff)
  # The idea here is that shows with few ratings either have unreliable stats
  # or are obsure or not valid matches
  filter(numRatings > 100) |> 
  # filter out shows that don't match years to eliminate partial matches
  # or books and movies with the same or similar titles
  filter(startYear == releaseYear) |> 
  # As an additional check, I will filter out entries that aren't classified
  # as TV shows
  filter(itemType.name == "TV Show") |>
  # Because TV series are queried by title (not IMDB ID), the same DTDD entry
  # can sometimes be returned as a match for two different IMDB titles that
  # have similar names. This creates duplicate DTDD IDs attached to different
  # IMDB titles, one of which is a false match.
  # To resolve this, for each duplicated DTDD ID, we compute the string edit
  # distance (Levenshtein distance) between the IMDB title and the name
  # DTDD has on record, and keep only the closest match. This assumes the
  # correct match will have the smaller distance between the two titles.
  group_by(id) |>
  mutate(dist = mapply(function(a, b) adist(a, b, ignore.case = TRUE)[1, 1], primaryTitle, name)) |>
  slice_min(dist, n = 1, with_ties = FALSE) |>
  ungroup() |>
  select(-dist)
           
```

```{r}
# Save the results
write_csv(results_all_tvseries, "data/imdb_tvseries_5000_dtdd_results.csv") # human readable
saveRDS(results_all_tvseries, "data/imdb_tvseries_5000_dtdd_results.rds") # for data processing
```

```{r}
# Fetch the series details for each series in the results using the id
# and save all the results into a new tibble

results_tvseries_details <- results_all_tvseries |>
  #slice_head(n = 10) |> # NOTE: This line is for debugging purposes only
  mutate(dtdd_details = map(id, fetch_dtdd_title_details)) |>
  unnest(dtdd_details) |>
  select(id, name, cleanName, 
         dog_yes_votes, dog_no_votes,
         cat_yes_votes, cat_no_votes,
         animal_yes_votes, animal_no_votes)
```

```{r}
# Save the series details
write_csv(results_tvseries_details, "data/imdb_tvseries_5000_dtdd_series_details.csv")
```

### Analyze the data

```{r}
# Create new columns in results_tvseries_details that calculate 
# the majority vote for each animal type and set to 1 if
# the majority vote is "yes" and 0 if the majority vote is "no".
# Set the value to 0 if there is no majority vote (i.e. if the yes and no votes are equal).
results_tvseries_details <- results_tvseries_details |>
  mutate(dog_death = case_when(
    dog_yes_votes > dog_no_votes ~ TRUE,
    TRUE ~ FALSE),
    cat_death = case_when(
      cat_yes_votes > cat_no_votes ~ TRUE,
      TRUE ~ FALSE),
    animal_death = case_when(
      animal_yes_votes > animal_no_votes ~ TRUE,
      TRUE ~ FALSE)
  )

saveRDS(results_tvseries_details, "data/imdb_tvseries_5000_dtdd_series_details.rds")
```

#### Compute death percentages

```{r}
# Compute percentages with CIs
compute_proportion_ci <- function(data, var) {
  n <- nrow(data)
  x <- sum(data[[var]], na.rm = TRUE)
  test <- prop.test(x, n)
  tibble(
    variable = var,
    pct = test$estimate * 100,
    ci_low = test$conf.int[1] * 100,
    ci_high = test$conf.int[2] * 100
  )
}

# Compute percentages and CIs for TV series deaths
tvseries_proportions <- map_df(
  c("dog_death", "cat_death", "animal_death"),
  ~ compute_proportion_ci(results_tvseries_details, .x)
)
print(tvseries_proportions)

# Test whether these differences in proportions are statistically significant

# Helper function to run prop.test comparing two proportions
compare_proportions <- function(data, var1, var2) {
  n <- nrow(data)
  x1 <- sum(data[[var1]], na.rm = TRUE)
  x2 <- sum(data[[var2]], na.rm = TRUE)
  prop.test(c(x1, x2), c(n, n))
}

# Compare animal vs dog and animal vs cat
test_animal_vs_dog <- compare_proportions(results_tvseries_details, "animal_death", "dog_death")
test_animal_vs_cat <- compare_proportions(results_tvseries_details, "animal_death", "cat_death")

print(test_animal_vs_dog)
print(test_animal_vs_cat)

```

#### Compute correlations between animal categories

```{r}
# Compute the correlation between a series having a dog death and
# having an animal death
cor.test(as.numeric(results_tvseries_details$dog_death),
         as.numeric(results_tvseries_details$animal_death),
         method = "pearson")

# Compute the correlation between a series having a cat death and
# having an animal death
cor.test(as.numeric(results_tvseries_details$cat_death),
         as.numeric(results_tvseries_details$animal_death),
         method = "pearson")

# Compute the correlation between a series having a dog death and
# having a cat death
cor.test(as.numeric(results_tvseries_details$dog_death),
         as.numeric(results_tvseries_details$cat_death),
         method = "pearson")
```

#### Analyze time trends

```{r}
# Plot deaths as a function of series start date

library(ggrepel)
library(patchwork)

p1 <- results_tvseries_details |>
  left_join(results_all_tvseries |> select(id, startYear), by = "id") |>
  filter(startYear >= 1999) |>
  pivot_longer(cols = c(dog_death, cat_death, animal_death),
               names_to = "death_type", values_to = "died") |>
  group_by(startYear, death_type) |>
  summarise(proportion = mean(died, na.rm = TRUE), .groups = "drop") |>
  mutate(death_type = recode(death_type,
    dog_death = "Dog",
    cat_death = "Cat",
    animal_death = "Other"
  )) |>
  group_by(death_type) |>
  mutate(is_last = startYear == max(startYear)) |>
  ungroup() |>
  ggplot(aes(x = startYear, y = proportion, 
             color = death_type, shape = death_type)) +
  geom_point() +
  geom_smooth(alpha = 0.3) +
  annotate("text", 
           x = 2025.5, 
           y = c(0.030, 0.12, 0.40), 
           label = c("Cat", "Dog", "Other"),
           color = "black",
           hjust = 0) +
  scale_color_brewer(palette = "Dark2") +
  scale_shape_manual(values = c(
    "Dog" = 16,
    "Cat" = 17,
    "Other" = 15
  )) +
  scale_y_continuous(labels = scales::percent) +
  scale_x_continuous(limits = c(1998.5, 2027)) +
  labs(x = "", y = "Proportion of\nTV series deaths") +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.title.y = element_text(angle = 0, vjust = 0.5),
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank()
  )

p2 <- results_tvseries_details |>
  left_join(results_all_tvseries |> select(id, startYear), by = "id") |>
  filter(startYear >= 1999) |>
  distinct(id, startYear) |>
  count(startYear) |>
  ggplot(aes(x = startYear, y = n)) +
  geom_col() +
  scale_x_continuous(limits = c(1998.5, 2027)) +
  labs(x = "Release year", y = "Number of\nTV series") +
  theme_minimal() +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    axis.title.y = element_text(angle = 0, vjust = 0.5),
  )

ggsave("figures/tv_trend_analysis.pdf", 
       plot = p1 / p2 + plot_layout(heights = c(4, 1), axes = "collect"),
       width = 7, 
       height = 4, 
       units = "in")
```

#### Analyze by genre

```{r}
# Analyze series by genre
# NOTE: I'm using the genres assigned by IMDB here and there are multiple
# genres for each series, so this will produce more data points than series

series_with_genre <- left_join(results_tvseries_details, results_all_tvseries |> select(id, genres), by = "id") |>
  separate_rows(genres, sep = ", ") 

# Pull only the top 15 most common genres
valid_genres_tv <- series_with_genre |> 
  count(genres) |> 
  slice_max(order_by = n, n = 15) |>
  pull(genres)

series_with_genre <- series_with_genre |> 
  filter(genres %in% valid_genres_tv)

# Calculate death rates by genre and CIs
genre_death_rates_tv <- series_with_genre |>
  group_by(genres) |>
  summarise(
    n = n(),
    dog_death_rate = mean(dog_death, na.rm = TRUE),
    cat_death_rate = mean(cat_death, na.rm = TRUE),
    animal_death_rate = mean(animal_death, na.rm = TRUE),
    dog_se = sqrt(dog_death_rate * (1 - dog_death_rate) / n),
    cat_se = sqrt(cat_death_rate * (1 - cat_death_rate) / n),
    animal_se = sqrt(animal_death_rate * (1 - animal_death_rate) / n)
  )

# Get genre order based on animal death rate
genre_order <- genre_death_rates_tv |>
  arrange(animal_death_rate) |>
  pull(genres)

label_order <- genre_death_rates_tv |>
  arrange(animal_death_rate) |>
  mutate(genre_label = paste0(genres, " (", n, ")")) |>
  pull(genre_label)

# Pivot to long format
genre_death_rates_long_tv <- genre_death_rates_tv |>
  pivot_longer(
    cols = c(dog_death_rate, cat_death_rate, animal_death_rate),
    names_to = "death_type",
    values_to = "proportion"
  ) |>
  mutate(
    genre_label = factor(paste0(genres, " (", n, ")"), levels = label_order),
    se = case_when(
      death_type == "dog_death_rate" ~ dog_se,
      death_type == "cat_death_rate" ~ cat_se,
      death_type == "animal_death_rate" ~ animal_se
    ),
    death_type = factor(death_type, 
                        levels = c("animal_death_rate", "dog_death_rate", "cat_death_rate"),
                        labels = c("Other", "Dog", "Cat"))
  )


# Movies plot
p_movies_genre <- ggplot(genre_death_rates_long_movies, aes(x = proportion, y = genre_label, 
                                                             color = death_type, shape = death_type)) +
  geom_point(size = 3) +
  # Clip the error bars at 0 to prevent them from going below b/c this doesn't make conceptual sense
  geom_errorbar(aes(xmin = pmax(proportion - 1.96 * se, 0), 
                  xmax = proportion + 1.96 * se),
              width = 0.2, alpha = 0.5, orientation = "y") +
  scale_color_manual(values = c("Cat" = "#1B9E77", "Dog" = "#D95F02", "Other" = "#7570B3")) +
  scale_shape_manual(values = c("Dog" = 16, "Cat" = 17, "Other" = 15)) +
  scale_x_continuous(labels = scales::percent) +
  labs(x = "Proportion of movies with a death", y = NULL, 
       color = NULL, shape = NULL, title = "Movies") +
  theme_minimal() +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5),
        plot.title.position = "panel",
        axis.title.x = element_blank(),
        panel.grid.minor = element_blank(),
        panel.grid.major.y = element_blank())

# TV series plot
p_tvseries_genre <- ggplot(genre_death_rates_long_tv, aes(x = proportion, y = genre_label,
                                                           color = death_type, shape = death_type)) +
  geom_point(size = 3) +
  # Clip the error bars at 0 to prevent them from going below b/c this doesn't make conceptual sense
  geom_errorbar(aes(xmin = pmax(proportion - 1.96 * se, 0), 
                  xmax = proportion + 1.96 * se),
              width = 0.2, alpha = 0.5, orientation = "y") +
  scale_color_manual(values = c("Cat" = "#1B9E77", "Dog" = "#D95F02", "Other" = "#7570B3")) +
  scale_shape_manual(values = c("Dog" = 16, "Cat" = 17, "Other" = 15)) +
  scale_x_continuous(labels = scales::percent) +
  labs(x = "Proportion of TV series with a death", y = NULL,
       color = NULL, shape = NULL, title = "TV Series") +
  theme_minimal() +
  theme(legend.position = c(0.84, 0.27),
        plot.title = element_text(hjust = 0.5),
        plot.title.position = "panel",
        legend.background = element_rect(fill = "white", linewidth = 0.5),
        panel.grid.minor = element_blank(),
        panel.grid.major.y = element_blank())

# Combine and save
ggsave("figures/genre_analysis.pdf", 
       plot = p_movies_genre / p_tvseries_genre,
       width = 7, 
       height = 6.5, 
       units = "in")
```

```{r}
# Plot the frequency of TV series and movies over time in the IMDB 5000 datasets

library(patchwork)

p_movies <- results_all_movies |>
  count(startYear) |>
  ggplot(aes(x = startYear, y = n)) +
  geom_col() +
  geom_vline(xintercept = 1999, color = "#E69F00", linetype = "dashed", linewidth = 0.5) +
  scale_x_continuous(limits = c(1914.5, 2026),
                     breaks = c(1920, 1940, 1960, 1980, 1999, 2020),
                     labels = c("1920", "1940", "1960", "1980", "1999", "2020")) +
  labs(x = "Year", y = NULL, title = "Movies") +
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank())

p_tvseries <- results_all_tvseries |>
  count(startYear) |>
  ggplot(aes(x = startYear, y = n)) +
  geom_col() +
  geom_vline(xintercept = 1999, color = "#E69F00", linetype = "dashed", linewidth = 0.5) +
  scale_x_continuous(limits = c(1915, 2026),
                     breaks = c(1920, 1940, 1960, 1980, 1999, 2020),
                     labels = c("1920", "1940", "1960", "1980", "1999", "2020")) +
  labs(x = "Year", y = NULL, title = "TV Series") +
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank())

p_movies + p_tvseries

ggsave("figures/movies_series_frequency_over_time.pdf", 
       plot = p_movies + p_tvseries,
       width = 7, 
       height = 3, 
       units = "in")
```